---
title: "Exploratory analysis"
author: "Jennah Gosciak"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
knit: (function(inputFile, encoding) { 
  rmarkdown::render(inputFile,
  encoding=encoding,
  output_file="00_data_prep.html")})
output:
  html_document:
    toc: true
    theme: paper
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = F, warning = F,
  fig.width = 6, fig.height = 10
)
```

```{r, setup2, include = F}
# Setup
library(renv)
library(tidyverse)
library(janitor)
library(magrittr)
library(eeptools)
library(assertr)
library(readr)
library(openxlsx)
library(DBI)
library(dplyr)
library(dbplyr)
library(RPostgres)
library(lubridate)
library(fs)
library(sf)
library(vroom)
library(leaflet)
library(haven)


# DB Connection
con <- dbConnect(Postgres(),
  user = "",
  password = rstudioapi::askForPassword(),
  host = "",
  port = ,
  dbname = "",
  bigint = "integer"
)

source("99_common.R")
out <- "03_Output"
inp <- "01_Input"
fig_path <- path("03_Output", "figs")
colors <- c("#512e92","#e03616","#758ecd", "#FFC857","#0b5351")
```

# Analysis: 1998-2021
Purpose: to create a cleaned file of NYU properties 1998-2021
Also runs other data-intensive tasks from the Furman database

```{r}
tbl(con, "rpad_raw") %>% 
  filter(year4 == 2000)
```

## Load earlier RPAD Files
```{r}
rpad_lt2002 <- st_read(con, query = "SELECT r.*, m.geom_poly,
                      m.geom_point FROM rpad_raw as r
                      LEFT JOIN mappluto_02b as m on coalesce(r.bbl, r.bble, r.bbl_condo) = m.bbl
                      WHERE r.year4 < 2002 AND
                      r.owner ~* 
                    '^(NYU)$|(\\sNYU,?\\s)|(NEW YORK UNIV.+)|^(NYU,?\\s)|(NY UNIV)|(POLYTECHNIC INST)'")
```
```{r}
# select only relevant vars
rpad_filt <- rpad_lt2002 %>% 
  select(c("bble", "bbl", "bbl_condo",  "cd_rpad",
            "cd", "bldgcl", "exmtcl",
           "owner", "hnum_lo", "hnum_hi", "str_name",
           "zip", "tot_unit", "res_unit", "bldgs",
           "lnd_area", "gr_sqft", "zoning", "yrb", "year4",
           "owner", "geom_poly", "geom_point",
           "hnum_lo", "hnum_hi", "str_name", "zip")) %>% 
  rename(bldgarea = gr_sqft,
         year = year4,
         ownername = owner,
         bldgclass = bldgcl) 

rpad_lt2002 %>% 
  st_set_geometry("geom_poly") %>% 
  st_transform(2263) %>% 
  select(-"geom_point") %>% 
  st_write(dsn = path(out, "rpad_nyu.geojson"), driver = "GeoJSON",
           append = F,
           delete_dsn = T)

rpad_lt2002 %>% 
  filter(st_is_empty("geom_poly")) %>% 
  nrow()
```

```{r}
# TEMP CODE
rpad_lt2002 <- st_read(path(out, "rpad_nyu.geojson"))
rpad_lt2002 %>% 
  tablist_qc(year4)


nyu_pluto <- st_read(path(out, "nyu_allyrs.geojson")) %>% 
  filter(year > 2001) %>% 
  as_tibble()

unmerged_rpad <- rpad_lt2002 %>% 
  as_tibble() %>% 
  anti_join(nyu_pluto, ., by = "bbl") %>% 
  select(c("bbl", "address", "ownername", "yearbuilt")) %>% 
  filter(yearbuilt < 2002) %>% 
  unique()
unmerged_rpad

bbl_list <- c("1009627501")
# 16 cooper sq, bought in 2004
# 120 east 12th st, the church

# 238 thompson street
# bought in 2009 from the archdiocese

# 60 5th avenue
# bought in 2010

# 433 EAST 30 STREET
# 1992 or 1998?

# 315 JAY ST WAS POLYTECHNIC

# 101 tech place
# was polytechnic

# 383 lafayette, lease in 2003
# deed in 2013

# 31 3rd avenue
# bought in 1984

# 467 3rd avenue
# bought in 1984

nyu_pluto %>% 
  filter(cd == 208) %>% 
  select(c("ownername", "address", "year", "yearbuilt", "bbl")) %>% 
  arrange(year)
```
## Try to identify some properties that aren't merging bc of missing data
```{r}
rpad_lt2002 <- rpad_lt2002 %>% 
  as_tibble() %>% 
  select(-"geometry")

copy_to(con, rpad_lt2002,
  name = "rpad_lt2002", overwrite = T,
  temporary = T
)

rpad_check <- st_read(con, query = "SELECT r.* FROM rpad_raw as r
                      INNER JOIN rpad_lt2002 as unmgd on coalesce(r.bbl, r.bble, r.bbl_condo) = unmgd.bbl
                      LEFT JOIN mappluto_02b as m on coalesce(r.bbl, r.bble, r.bbl_condo) = m.bbl
                      WHERE r.year4 < 2002 and r.year4 > 1996")

rpad_check %>% 
  select(c("owner", "bbl", "str_name"))

rpad_check %>% 
  st_write(dsn = path(out, "rpad_unmerged.geojson"),
           append = F,
           dsn_delete = T)
```
```{r}
rpad_unmerged <- st_read(path(out, "rpad_unmerged.geojson")) %>% 
  filter(is.na(owner)) %>% 
  #filter(is.na(owner) & str_sub(exmtcl, 1, 1) == "X") %>% 
  mutate(l_acre = as.character(l_acre),
         bld_var = as.character(bld_var),
         ex_chgdt = as.character(ex_chgdt))

rpad_unmerged %>% 
  tablist_qc(year4)

rpad_unmerged %>% 
  tablist_qc(owner)

rpad_final <- rpad_lt2002 %>% 
  as_tibble() %>% 
  bind_rows(rpad_unmerged %>% 
              as_tibble()) %>% 
  unique()  %>% 
  rename(bldgarea = gr_sqft,
         year = year4,
         ownername = owner,
         bldgclass = bldgcl) %>% 
  select(-c("bble", "ease")) %>% 
  group_by(bbl, year) %>% 
  mutate(drop_row = case_when(n() > 1 & !is.na(ownername) ~ 0,
                              n() == 1 ~ 0,
                              TRUE ~ 1)) %>% 
  filter(drop_row == 0) 

rpad_final %>% 
  as_tibble() %>% 
  verify(!is.na(year))

rpad_final %>% 
  tablist_qc(year)

rpad_final %>% 
  as_tibble() %>% 
  verify_isid(c("bbl", "year"))

## TEMP CODE combine
nyu_upd <- bind_rows(as_tibble(nyu_pluto), rpad_final) %>% 
  st_as_sf()

nyu_upd %>% 
  tablist_qc(year)

rpad_final %>% 
  verify(!is.na(year)) %>% 
  tablist_qc(year)

st_write(st_as_sf(rpad_final) %>% 
           select(-"drop_row"), dsn = path(out, "rpad_upd.geojson"),
         append = F,
         dsn_delete = T)

st_write(nyu_upd, dsn = path(out, "nyu_allyrs_upd.geojson"),
         append = F,
         dsn_delete = T)
```



## Load 2008 PLUTO File (not on Furman Database)
```{r}
boros <- c("Manhattan", "Brooklyn", "Queens", "Bronx",
           "Staten_Island")
abbs <- c("MN", "BK", "QN", "BX", "SI")
# load 2008 data
pluto2008 <- map2_dfr(boros, abbs,
                      ~st_read(path(inp, "mappluto_08b", 
                                    "mappluto_08b", .x, paste0(.y, "MapPLUTO.shp"))))  %>% 
  mutate(year = 2008)

# convert lowercase
lowernames <- names(pluto2008) %>% 
  str_to_lower()
names(pluto2008) <- lowernames

# pull out assessed value
pluto2008_assessed_value <- pluto2008 %>% 
  as_tibble() %>% 
    rename_at(vars(starts_with("assesstot")), ~"assessed_value") %>% 
    rename_at(vars(starts_with("cd")), ~"cd") %>%
    rename_at(vars(starts_with("floorarea")), ~"bldgarea") %>% 
    rename_at(vars(starts_with("landuse")), ~"landuse") %>% 
    # filter for mixed commercial or commercial and office buildings
    filter(landuse == "04" | landuse == "05") %>% 
    group_by(cd) %>% 
    mutate(value_sqft = if_else(bldgarea > 0, assessed_value / bldgarea, NA_real_)) %>% 
    summarize(avg_value_sqft = round(mean(value_sqft, na.rm = T), 2)) %>% 
    mutate(year = 2008)
pluto2008_assessed_value
```
```{r}
# filter out unncessary vars
pluto2008_nyu <- pluto2008 %>% 
    filter(str_detect(ownername, '^(NYU)$|(\\sNYU,?\\s)|(NEW YORK UNIV.+)|^(NYU,?\\s)|(NY UNIV)|(POLYTECHNIC INST)'),
         ownername != 'NYU, SOON KEAT') %>% 
    mutate(geom_poly = geometry,
           geom_point = st_centroid(geometry)) %>% 
    select("ownername", matches("area"),
           matches("cd"),
           matches("assess"),
           matches("yearbuilt"),
           matches("address"), "bbl",
           "year",
           "geom_poly",
           "geom_point") %>% 
  mutate(bbl = as.character(bbl)) %>% 
  st_set_geometry("geom_poly") %>% 
  select(-"geometry") 

pluto2008_nyu

nrow(pluto2008_nyu)
```

## Load all other PLUTO Files
```{r}
check_extract <- function(con, tblname) {
  df1 <- tbl(con, tblname) %>%
    mutate(ownername = str_replace_all(ownername, "[:punct:]", "")) %>%
    filter(str_detect(ownername, "NEW YORK UNI")) %>%
    select(ownername) %>%
    as_tibble() %>%
    unique()

  df2 <- tbl(con, tblname) %>%
    mutate(ownername = str_replace_all(ownername, "[:punct:]", "")) %>%
    filter(str_detect(ownername, "NYU")) %>%
    select(ownername) %>%
    as_tibble() %>%
    unique()

  bind_rows(df1, df2)
}

extract_nyu <- function(con, tblname, yr, out_path = out) {
  print(paste0("Loading table for: ", tblname))
  
  nyu_pluto <- st_read(con, query = paste0(
    "SELECT * FROM ", tblname,
    " WHERE (ownername ~* '^(NYU)$|(\\sNYU,?\\s)|(NEW YORK UNIV.+)|^(NYU,?\\s)|(NY UNIV)|(POLYTECHNIC INST)'
                  AND ownername != 'NYU, SOON KEAT') OR bbl = '1005330001'"
  )) %>%
    mutate(year = yr)
  
  lower_names <- str_to_lower(names(nyu_pluto))
  names(nyu_pluto) <- lower_names
  
  nyu_pluto <- nyu_pluto %>% 
    select("ownername", matches("area"),
           matches("cd"),
           matches("assess"),
           matches("yearbuilt"),
           matches("address"),
           matches("bldg"), 
           matches("land"), "bbl",
           "year",
           "geom_poly",
           "geom_point")
    
  nyu_pluto %>%
    nrow()

  nyu_pluto %>%
    tablist_qc(ownername)

  return(nyu_pluto)
}
```
```{r}
tbls <- dbListTables(con)
tbls[str_detect(tbls, "pluto") & str_detect(tbls, "09")]

tblnames <- c(
  "mappluto_02b", "mappluto_03c", "mappluto_04c", "mappluto_05d",
  "mappluto_06c", "mappluto_07c", "mappluto_09v2",
  "mappluto_10v2", "mappluto_11v2",
  "mappluto_12v2", "mappluto_13v2", "mappluto_14v2", "mappluto_15v1",
  "mappluto_16v2", "mappluto_17v1_1", "mappluto_18v2", "mappluto_19v2",
  "mappluto_20v1", "mappluto_21v3"
)

years <- c(2002:2007, 2009:2021)
nyu_pluto_full <- map2(tblnames, years, ~ extract_nyu(con, .x, .y))
```
```{r}
nyu_pluto_app <- bind_rows(nyu_pluto_full, pluto2008_nyu, rpad_filt)
nyu_pluto_app

nyu_21 <- extract_nyu(con, "mappluto_21v3", 2021)
nyu_21
```

Export output
```{r}
drop_vars <- c("shape_area",
               "strgearea",
               "garagearea",
               "retailarea",
               "areasource",
               "otherarea",
               "landmark",
               "healtharea",
               "officearea",
               "ind_area",
               "comarea",
               "landmkdate",
               "factryarea",
               "areasource")

nyu_21 %>%
  st_write(dsn = path(out, "nyu_21.geojson"), driver = "GeoJSON",
           append = F,
           delete_dsn = T)

nyu_pluto_app %>%
  st_set_geometry("geom_poly") %>% 
  st_transform(2263) %>% 
  select(-"geom_point") %>% 
  select(-any_of(drop_vars)) %>% 
  st_write(dsn = path(out, "nyu_allyrs.geojson"), driver = "GeoJSON",
           append = F,
           delete_dsn = T)

nyu_pluto_app %>%
  st_set_geometry("geom_point") %>% 
  st_transform(2263) %>% 
  select(-"geom_poly") %>% 
  select(-any_of(drop_vars)) %>% 
  st_write(dsn = path(out, "nyu_allyrs_pts.geojson"), driver = "GeoJSON",
           append = F,
           delete_dsn = T)
```

```{r}
check_names <- map_dfr(tblnames, ~ check_extract(con, .)) %>%
  unique()

check_names %>%
  write.xlsx(path(out, "check_names.xlsx"), overwrite = T)

check_names %>%
  anti_join(nyu_pluto_app, by = "ownername") %>%
  write.xlsx(path(out, "check_unmg_names.xlsx"), overwrite = T)
```
```{r}
bbl_unmgd <- nyu_pluto_app %>%
  as_tibble() %>%
  select(c("bbl")) %>%
  unique()

# duplicate list of all owner names, add year variable
# use this to merge back onto the original file
bbl_comb <- map_dfr(years, ~ mutate(bbl_unmgd, year = .))

# use anti join to see what's not merghing
bbl_unmgd <- bbl_comb %>%
  anti_join(nyu_pluto_app, by = c("bbl", "year"))

bbl_unmgd %>%
  group_by(year) %>%
  summarize(num_unmgd = n_distinct(bbl))

copy_to(con, bbl_unmgd,
  name = "bbl_unmgd", overwrite = T,
  temporary = T
)

check_pluto <- function(con, tbl, yr) {
  tbl(con, tbl) %>%
    inner_join(tbl(con, "bbl_unmgd") %>%
      filter(year == yr), by = "bbl") %>%
    select(c("ownername", "bbl", "year")) %>%
    as_tibble()
}

full_check <- map2_dfr(tblnames, years, ~ check_pluto(con, .x, .y))
full_check %>%
  write.xlsx(path(out, "unmgd_by_year.xlsx"), overwrite = T)
```

## Calculate average assessed value for all commercial buildings in a CD
```{r}
pluto_calc_assess_value <- function(con, tblname, yr) {
  tbl(con, tblname) %>% 
    rename_at(vars(starts_with("assesstot")), ~"assessed_value") %>% 
    rename_at(vars(starts_with("cd")), ~"cd") %>%
    rename_at(vars(starts_with("floorarea")), ~"bldgarea") %>% 
    rename_at(vars(starts_with("landuse")), ~"landuse") %>% 
    # filter for mixed commercial or commercial and office buildings
    filter(landuse == "04" | landuse == "05") %>% 
    group_by(cd) %>% 
    mutate(value_sqft = if_else(bldgarea > 0, assessed_value / bldgarea, NA_real_)) %>% 
    summarize(avg_value_sqft = round(mean(value_sqft, na.rm = T), 2)) %>% 
    mutate(year = yr) %>% 
    as_tibble() 
}

# calc average assessed value in commercial and mixed commercial buildings
# by community district
# use 2002 values for 1998-2001, since we don't have data for those years
assessed_value_avg <- map2_dfr(c(rep(tblnames[1], 4),
                                 tblnames), c(1998:2001, years), ~pluto_calc_assess_value(con, .x, .y)) 
assessed_value_avg_2008 <- bind_rows(assessed_value_avg, pluto2008_assessed_value)

assessed_value_avg_2008 %>% 
  saveRDS(path(out, "assess_value_averages.RDS"))

assessed_value_avg_2008 %>% 
  write_csv(path(out, "assess_value_averages.csv"))
```
