---
title: "Analysis of RPAD and PLUTO Data"
author: "Jennah Gosciak"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
knit: (function(inputFile, encoding) { 
  rmarkdown::render(inputFile,
  encoding=encoding,
  output_file="01_analysis_2001_2021.html")})
output:
  html_document:
    toc: true
    theme: paper
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = F, warning = F,
  fig.width = 6, fig.height = 10
)
```

```{r, setup2, include = F}
# Setup
library(renv)
library(tidyverse)
library(janitor)
library(magrittr)
library(eeptools)
library(assertr)
library(readr)
library(openxlsx)
library(DBI)
library(dplyr)
library(dbplyr)
library(RPostgres)
library(lubridate)
library(fs)
library(sf)
library(vroom)
library(leaflet)
library(haven)
library(geojsonsf)
library(s2)
library(geoclient)
source("99_common.R")

out <- "03_Output"
inp <- "01_Input"
fig_path <- path("03_Output", "figs")
web_path <- path("04_WebMap", "data")
spatial_path <- path("03_Output", "spatial_files")
colors <- c("#512e92","#e03616","#758ecd", "#FFC857","#0b5351")
alt_purple <- "#9057FF"

# DB Connection
con <- dbConnect(Postgres(),
  user = "",
  password = rstudioapi::askForPassword(),
  host = "",
  port = ,
  dbname = "",
  bigint = "integer"
)
```

# Analysis: 2001-2021
* Note: there were too many missing data issues for 1998-2000, so I am only using 2001 from the RPAD data.

## Load cpi conversion factors
* From the [Bureau of Labor Statistics](https://www.bls.gov/cpi/research-series/r-cpi-u-rs-home.htm)
```{r, exp pluto}
# load cpi from bls
# method from: https://qrc.depaul.edu/oelguntillman/Winter12/Notes/7%20-%20CPI%20Notes.htm
cpi <- readxl::read_xlsx(path(inp, "r-cpi-u-rs-allitems.xlsx"),
                         skip = 5)
cpi
cpi_2020 <- cpi %>% 
  filter(YEAR == 2020) %>% 
  pull(AVG)
conversion_factors <- cpi %>% 
  filter(YEAR != 1977) %>% 
  mutate(conv = cpi_2020/AVG) %>% 
  rename(year = YEAR) %>% 
  select(year, conv)
conversion_factors
```

## Load data
* Points
* Polygons
* Just 2021
```{r}
# read in point data
nyu_pts <- st_read(path(out, "nyu_allyrs_pts.geojson")) %>% 
  filter(year >= 2001) %>% 
  clean_base_file()

nyu_poly <- st_read(path(out, "nyu_allyrs.geojson")) %>% 
  filter(year >= 2001) %>% 
  clean_base_file() %>% 
  group_by(bbl) %>% 
  fill(address)

nyu_21 <- st_read(path(out, "nyu_21.geojson")) %>% 
  filter(!(ownername %in% c("YOON, SOOK NYU", "YOON, SUK NYU", "ARA HOLDINGS OF NYU L",
                            "ARA HOLDINGS OF NYU LLC")))

rpad_allyrs <- readRDS(path(out, "rpad_allyrs_pts.RDS")) %>% 
  filter(!(str_trim(ownername) %in% c("YOON, SOOK NYU", "YOON, SUK NYU","NYU, SOON KEAT",
                            "ARA HOLDINGS OF NYU L",
                            "ARA HOLDINGS OF NYU LLC",
                            "REDUX NYU  LLC"))) %>% 
  unique()

rpad_allyrs %>% 
  tablist_qc(ownername)

# checking uniqueness
rpad_allyrs %>% 
  as_tibble() %>% 
  verify_isid(c("bbl", "year"))
```
```{r}
rpad_geo <- rpad_allyrs %>% 
  mutate(across(c("hnum_lo", "hnum_hi"), ~str_replace_all(., "O", "0"))) %>% 
    mutate(hnum_lo = if_else(is.na(hnum_lo), hnum_hi, hnum_lo),
           address_form = case_when(!is.na(str_name) ~ str_c(paste(as.numeric(hnum_lo), 
                                                                    str_to_sentence(str_trim(str_name))), 
                                                                   "New York", "New York", zip, sep = ", "),
                                    TRUE ~ NA_character_),
           address_form = case_when(hnum_lo == "00000001 1/2"  | hnum_lo == "1 1/2" ~ 
                                      "1 1/2 Waverly Place, New York, New York 10003",
                                    address_form == "425 Main street, New York, New York, 0" ~ 
                                      "425 Main street , New York, New York, 10044",
                                    address_form == "NA West tremont avenue, New York, New York, 0" ~ 
                                      "500 Mac Cracken Avenue, New York, New York, 10453",
                                    address_form == "NA River road, New York, New York, 0" ~
                                      "2 River Avenue, New York, New York, 10463",
                                    TRUE ~ address_form),
           ) %>% 
  geo_search_data(address_form) %>% 
  mutate(bbl_geo = bbl) %>% 
  select(matches("input"), "no_results", "bbl_geo") %>%
  cbind(rpad_allyrs) %>% 
  verify(!is.na(no_results)) 
rpad_geo %>% 
  tablist_qc(no_results)

## merge to pluto 21
copy_to(con, rpad_geo,
  name = "rpad_geo", overwrite = T,
  temporary = T
)

rpad_geo_merge <- st_read(con, query = "SELECT r.*, m.geom_point FROM rpad_geo as r
                          LEFT JOIN mappluto_21v3 as m on r.bbl_geo = m.bbl")  %>% 
  st_transform(4326) %>% 
  as_tibble()

rpad_geo_merge

# replace missing values
# manually for one address
replace_address <- tibble(address_form = c("158 East 34 street, New York, New York, 10016",
                                      "156 East 34 street, New York, New York, 10016"),
                 lat = c(-73.97900442578594, -73.97900442578594),
                 lon = c(40.74592051070297, 40.74592051070297)) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% 
  as_tibble() %>% 
  rename(geom_new = geometry) %>% 
  mutate(geom_new = st_sfc(geom_new))

replace_address

rpad_geo_fix <- rpad_geo_merge %>%
  rename(address_form = input_location) %>%
  as_tibble() %>% 
  left_join(replace_address, by = "address_form") %>% 
  mutate(geom_point = ifelse(st_is_empty(geom_point), geom_new, geom_point)) %>% 
  select(-"geom_new") %>% 
  st_set_geometry("geom_point") %>% 
  group_by(bbl_condo, year) %>% 
  #if more than one per buildings etc.
  mutate(dupl_pts = ifelse(n() > 1, 1, 0)) %>% 
  ungroup() %>% 
  mutate(geom_point = ifelse(dupl_pts == 1, st_jitter(geom_point, amount = 0.0002), geom_point)) %>% 
  as_tibble() %>% 
  st_as_sf()
rpad_geo_fix

rpad_geo_fix %>% 
  filter(st_is_empty(geom_point)) %>% 
  nrow() %>% 
  equals(0) %>% 
  stopifnot()

rpad_geo_fix %>% 
  st_write(path(web_path, "rpad_geo.geojson"),
           append = F,
           dsn_delete = T)

rpad_geo_fix %>% 
  filter(year == 2021) %>% 
  st_write(path(web_path, "rpad_geo_21.geojson"),
           append = F,
           dsn_delete = T)

# check that there are no duplicates
# random example I had found in the map data
rpad_geo_fix %>%
  as_tibble() %>% 
  verify_isid(c("bbl", "year"))

rpad_geo_fix %>% 
  select(c("bbl", "year"), everything()) %>% 
  filter(year == 2021, cd == 315)

rpad_geo_fix %>% 
  select(c("bbl", "year"), everything()) %>% 
  filter(year == 2021, cd == 315) %>% 
  ggplot() + 
  geom_sf()
```
## Loading background shapefiles for context maps
```{r}
url_boro <- "https://services5.arcgis.com/GfwWNkhOj9bNBqoJ/arcgis/rest/services/NYC_Borough_Boundary/FeatureServer/0/query?where=1=1&outFields=*&outSR=4326&f=pgeojson"
url_nhood <- "https://services5.arcgis.com/GfwWNkhOj9bNBqoJ/arcgis/rest/services/NYC_Neighborhood_Tabulation_Areas_2020/FeatureServer/0/query?where=1=1&outFields=*&outSR=4326&f=pgeojson"
#url_streets <- "https://data.cityofnewyork.us/resource/m6hn-ddkx.geojson?$limit=200000"
url_roads <- "https://data.cityofnewyork.us/resource/gdww-crzy.geojson"
url_cds <- "https://data.cityofnewyork.us/resource/jp9i-3b7y.geojson?$limit=200000"
url_parks <- "https://data.cityofnewyork.us/resource/y6ja-fw4f.geojson?$limit=200000"
# load boros
boro <- st_read(url_boro) %>% 
  st_transform(2263)

# load major streets
# endpoint is not loading--temporarily commented out
# streets <- st_read(url_streets) %>%
#   st_transform(2263)
# nrow(streets)

# load cds
cds <- st_read(url_cds)  %>% 
  st_transform(2263)

# load parks
parks <- st_read(url_parks) %>%
  st_transform(2263)

# load nhood
nhood <- st_read(url_nhood)  %>% 
  st_transform(2263)
nyu_poly
```

## NYU in 2021
Starts with general background statistics

```{r}
# Identify list of community districts
intersecting_cds <- nyu_pts %>% 
  filter(year == 2021) %>% 
  st_join(cds, join = st_intersects) %>% 
  filter(!is.na(boro_cd)) %>% 
  pull(boro_cd) %>% 
  unique()

# checking geometry works
# example of plot
nyu_poly %>% 
  filter(year == 2021) %>% 
  ggplot() +
  geom_sf(data = boro, color = "gray", fill = NA) +
  geom_sf(data = cds, color = "black", fill = NA, alpha = 0.5) +
  geom_sf(fill = colors[1], color = NA) +
  theme_void()

nyu_poly %>% 
  filter(year == 2021) %>% 
  as_tibble() %>% 
  pull(cd) %>% 
  unique()

nyu_poly %>% 
  filter(year == 2021) %>% 
  as_tibble() %>% 
  arrange(cd) %>% 
  select(c("cd", "bbl", "ownername", "address"), everything()) %>% 
  verify(!is.na(bldgarea))

nyu_poly %>% 
  filter(year == 2021) %>% 
  as_tibble() %>% 
  arrange(cd) %>% 
  select(c("cd", "bbl", "ownername", "address"), everything()) %>% 
  verify(!is.na(bldgarea)) %>% 
  group_by(cd) %>% 
  summarize(total_bldarea = sum(bldgarea),
            total_bld = n_distinct(bbl))
nyu_poly %>% 
  filter(year == 2021) %>% 
  as_tibble() %>% 
  arrange(cd) %>% 
  select(c("cd", "bbl", "ownername", "address"), everything()) %>% 
  verify(!is.na(bldgarea)) %>% 
  filter(address == "28 BETHUNE STREET" |
         address == "63 CHARLES STREET")

nyu_poly %>% 
  filter(year == 2021) %>% 
  as_tibble() %>% 
  arrange(cd) %>% 
  select(c("cd", "bbl", "ownername", "address"), everything()) %>% 
  verify(!is.na(bldgarea)) %>% 
  filter(cd == 307)

total_lotarea <- tbl(con, "mappluto_21v3") %>%
  filter(bldgclass %in% c("W5", "W6")) %>%
  summarize(total_lotarea = sum(lotarea, na.rm = T),
            total_bldgarea = sum(bldgarea, na.rm = T)) %>%
  as_tibble()

total_lotarea
nyu_21 %>% 
  as_tibble() %>% 
  summarize(nyu_lotarea = sum(lotarea, na.rm = T),
            nyu_bldgarea = sum(bldgarea, na.rm = T))
  
nyu_21 %>%
  # these are university building classes
  filter(bldgclass %in% c("W5", "W6")) %>%
  as_tibble() %>%
  summarize(nyu_lotarea = sum(lotarea, na.rm = T),
            nyu_bldgarea = sum(bldgarea, na.rm = T)) %>%
  cbind(total_lotarea) %>%
  mutate(nyu_lotshare = nyu_lotarea * 100 / total_lotarea,
         nyu_bldgshare = nyu_bldgarea * 100 / total_bldgarea)

nyu_21 %>% 
  # this is mixed use, commercial, and residential
  filter(landuse %in% c("01", "02", "03", "04")) %>% 
  as_tibble() %>% 
  summarize(nyu_lotarea = sum(lotarea, na.rm = T),
            nyu_bldgarea = sum(bldgarea, na.rm = T))
nyu_21 %>% 
  # this is building class for healthcare
  filter(str_sub(bldgclass, 1, 1) == "I") %>% 
  as_tibble() %>% 
  summarize(nyu_lotarea = sum(lotarea, na.rm = T),
            nyu_bldgarea = sum(bldgarea, na.rm = T))

nyu_21 %>% 
  # these are university building classes
  filter(bldgclass %in% c("W5", "W6")) %>% 
  as_tibble() %>% 
  pull(landuse) %>% 
  unique()
nyu_21 %>%
  summarize(assesstot = sum(assesstot, na.rm = T))
```


Output all cds with nyu property in 2021

```{r}
# walk(intersecting_cds, ~expt_cd(cds, nyu_poly, ., 2021, spatial_path, fig_path, streets, parks))
```

Produce context map
```{r}
nyu_21 %>%
  ggplot() +
  geom_sf(data = parks,
            fill = "#777777", color = NA) +
  geom_sf(fill = "#9057FF", color = NA) +
  geom_sf(data = cds, color = "#777777", fill = NA, size = 0.1) +
  theme_void() +
  theme(
    panel.border = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    plot.background = element_rect(fill = '#A5A5A5', color = NA),
    panel.background = element_rect(fill = '#A5A5A5', color = NA)
  )

ggsave(path(fig_path, "nyu_21_map.pdf"), height = 12.5, width = 24)
```

Context map of Washington square park (just Manhattan)
```{r}
ggplot(nyu_21) +
  geom_sf(data = boro %>% 
            filter(BoroCode == 1), fill = NA, color = "white", size = 0.05) +
  geom_sf(data = parks %>%
            filter(str_sub(parknum, 1, 1) == "M"), fill = "#777777", color = NA) +
  geom_sf(fill = "#9057FF", color = NA) +
  theme_void() +
  theme(plot.background = element_rect(fill = '#424242'))
ggsave(path(fig_path, "mn_map_nyu.pdf"), width = 9, height = 9)
```

## Produce summmary tables for 2021
```{r}
tbl1 <- nyu_21 %>% 
  summarize(bldgarea_nyu = sum(bldgarea, na.rm = T),
            lotarea_nyu = sum(lotarea, na.rm = T),
            assessval_nyu = sum(assesstot, na.rm = T),
            bbl_nyu = n_distinct(bbl))
# load 2021 pluto
tbl1 <- tbl(con, "mappluto_21v3") %>%
  filter(!(ownertype %in% c("C", "O"))) %>%
  summarize(bldgarea_total = sum(bldgarea, na.rm = T),
            lotarea_total = sum(lotarea, na.rm = T),
            assessval_tot = sum(assesstot, na.rm = T),
            bbl_tot = n_distinct(bbl)) %>%
  as_tibble() %>%
  cbind(tbl1) %>%
  mutate(bldgarea_per = round(bldgarea_nyu * 100/ bldgarea_total, 2),
         lotarea_per = round(lotarea_nyu * 100 / bldgarea_total, 2),
         assessval_per = round(assessval_nyu * 100 / assessval_tot, 2),
         bbl_per = round(bbl_nyu * 100 / bbl_tot, 2))

tbl1
tbl1 %>%
  select(-c("bldgarea_total", "lotarea_total", "assessval_tot", "bbl_tot", "geometry")) %>%
  pivot_longer(everything()) %>%
  mutate(type = if_else(str_detect(name, "nyu"), "nyu", "per"),
         name = str_replace_all(name, "(_nyu)|(_per)", "")) %>%
  pivot_wider(name, names_from = "type", values_from = "value") %>%
  ggplot(aes(reorder(name, desc(per)), per)) +
  geom_col(fill = colors[1]) +
  geom_text(aes(label = nyu), vjust = -1) +
  theme_classic() +
  scale_y_continuous(limits = c(0, 3), breaks = c(0, 1, 2, 3)) +
  labs(x = "",
         y = "Percent of non-city owned properties")
ggsave(path(fig_path, "nyu_prop_summ.pdf"))

tbl(con, "mappluto_21v3") %>% 
  filter(ownername == "NATIONAL PARK SERVICE")
```

```{r}
# summary of top owners for 2021
owner_summ <- tbl(con, "mappluto_21v3") %>%
  mutate(ownername = str_trim(str_remove_all(ownername, "[\\.,-:']")),
         ownername = str_replace_all(ownername, "(CORP\\s|CORP$)", "CORPORATION")) %>%
  # filter out city and other public, state, or federal
  filter(!(ownertype %in% c("C", "O"))) %>%
  # filter out unknown and other city/government orgs
  filter(!(ownername %in% c("UNAVAILABLE OWNER", "NAME NOT ON FILE",
                            "NATIONAL PARK SERVICE",
                            "UNITED STATES OF AMERICA",
                            "PORT AUTHORITY OF NY AND NJ",
                            "UNITED STATES POSTAL SERVICE",
                            "MTA  LIRR",
                            "NYS DEPARTMENT OF TRANSPORTATION",
                            "MTA  STATEN ISLAND RAILWAY",
                            "HOUSING TRUST FUND CORPORATION",
                            "NYC DEPARTMENT OF ENVIRONMENTAL PROTECTION",
                            "NYS DIV OF LANDS AND FORESTS",
                            "DORMITORY AUTHORITY OF THE STATE OF NEW YORK",
                            "NYC TRANSIT AUTHORITY",
                            "NYS DEPARTMENT OF ENVIRONMENTAL CONSERVATION",
                            "CONSOLIDATED EDISON COOF NYINC",
                            "NYC DEPARTMENT OF TRANSPORTATION",
                            "CITY OF NEW YORK",
                            "MTA METRO NORTH",
                            "NYC HOUSING AUTHORITY",
                            "METROPOLITAN TRANSPORTATION AUTHORITY",
                            "HOUSING TRUST FUND CORP",
                            "USPS",
                            "NYS OFFICE OF GENERAL SERVICES",
                            "MTA  METRO NORTH",
                            "NYC HOUSING DEVELOPMENT CORPORATION"
                            ))) %>%
  group_by(ownername) %>%
  summarize(bldgarea_total = sum(bldgarea, na.rm = T),
            lotarea_total = sum(lotarea, na.rm = T),
            assessval_tot = sum(assesstot, na.rm = T),
            bbl_tot = n_distinct(bbl)) %>%
  ungroup() %>%
  arrange(desc(bldgarea_total))  %>%
  as_tibble()

# includes the city
owner_summ_city <- tbl(con, "mappluto_21v3") %>%
  filter(!(ownername %in% c("UNAVAILABLE OWNER", "NAME NOT ON FILE"))) %>%
  group_by(ownername) %>%
  summarize(bldgarea_total = sum(bldgarea, na.rm = T),
            lotarea_total = sum(lotarea, na.rm = T),
            assessval_tot = sum(assesstot, na.rm = T),
            bbl_tot = n_distinct(bbl)) %>%
  ungroup() %>%
  arrange(desc(bldgarea_total)) %>%
  as_tibble()

#Top ownership tables
owner_summ
# includes the city
owner_summ_city
owner_summ %>%
  arrange(desc(bbl_tot))
```

## Graph of top ownership in NYC in 2021
```{r, fig.width = 10}
owner_summ %>% 
  head(10) %>% 
  mutate(owner_form = str_to_sentence(ownername),
         bldgarea_total = bldgarea_total / 1000) %>% 
  ggplot(aes(reorder(owner_form, bldgarea_total), bldgarea_total))+
  geom_col(fill = colors[1]) +
  geom_text(aes(label = round(bldgarea_total)), hjust = 1.2, color = "white") +
  coord_flip() +
  theme_classic() +
  scale_y_continuous(limits = c(0, 18000)) +
  labs(y = "Building sq. ft.",
       x = "")
ggsave(path(fig_path, "ranking_bldsqft.pdf"))
owner_summ %>%
  arrange(desc(bbl_tot)) %>% 
  head(10) %>% 
  mutate(owner_form = str_to_sentence(ownername)) %>% 
  ggplot(aes(reorder(owner_form, bbl_tot), bbl_tot))+
  geom_col(fill = colors[1]) +
  geom_text(aes(label = bbl_tot), hjust = 1.2, color = "white") +
  coord_flip() +
  theme_classic() +
  scale_y_continuous(limits = c(0, 180)) +
  labs(y = "Number of buildings",
       x = "")
ggsave(path(fig_path, "ranking_nbld.pdf"))
```

## Tax and assessment calculations
```{r}
# Class 4: All commercial and industrial properties, such as office, retail, factory buildings and all other properties not 
# included in tax classes 1, 2 or 3.
## Load tax rate
tax_rate <- read_csv(path(inp, "Property_Tax_Rates_by_Tax_Class.csv")) %>% 
  select(c("YEAR", "CLASS 4")) %>% 
  rename(year = YEAR) %>% 
  filter(!is.na(year)) %>% 
  mutate(rate = as.numeric(str_replace_all(`CLASS 4`, "%", ""))/100)
tax_rate
```

```{r}
# load assessed_value_avg
assessed_value_avg <- readRDS(path(out, "assess_value_averages.RDS"))

nyu_assess_value <- nyu_poly %>%
  mutate(borough = str_sub(cd, 1, 1),
           cd = as.numeric(cd)) %>% 
  left_join(conversion_factors, by = "year") %>% 
  left_join(assessed_value_avg, by = c("cd", "year")) %>% 
  left_join(tax_rate, by = "year") %>% 
  rowwise() %>%
  mutate(assessed_value_orig = max(assesstotl, 
                                   assesstota, 
                                   assesstot, na.rm = T)) %>% 
  ungroup() %>% 
  # estimate taxes
  mutate(taxes_calc = rate * assessed_value_orig) %>% 
  group_by(year) %>% 
  mutate(taxes_yr = sum(taxes_calc, na.rm = T)) %>% 
  ungroup() %>% 
  # re-estimate assessed_value (doesn't produce better estimate)
  mutate(bldgarea = case_when(is.na(bldgarea) | bldgarea == 0 ~ floorarea,
                              TRUE ~ as.double(bldgarea))) %>%
  # estimate assessed value -- use lotarea if still missing bldgarea
  mutate(assessed_value = if_else(is.na(bldgarea) | bldgarea == 0, 
                                  lotarea * avg_value_sqft,
                                  bldgarea * avg_value_sqft)) %>% 
  mutate(borough = recode(borough, "3" = "Brooklyn",
                          "1" = "Manhattan",
                          "2" = "Bronx", 
                          "5" = "Staten Island",
                          "4" = "Queens")) %>%
  # adjust for inflation
  mutate(assessed_adj = if_else(year < 2021,
                                round(assessed_value * conv),
                                assessed_value),
         assessed_adj_orig = if_else(year < 2021,
                                round(assessed_value_orig * conv),
                                assessed_value_orig),
         taxes_calc = if_else(year < 2021,
                              round(taxes_calc * conv),
                              taxes_calc),
         taxes_yr = if_else(year < 2021,
                            round(taxes_yr * conv),
                            taxes_yr))

nyu_assess_geo <- nyu_assess_value  %>% 
  st_transform(4326) %>%
  mutate(bldgarea = case_when(is.na(bldgarea) ~ floorarea,
                              TRUE ~ bldgarea)) %>% 
  select(
    "ownername", "year", "assessed_adj",
    "assessed_adj_orig",
    matches("area"),
    "cd",
    "address_form", "bbl", "borough",
    "taxes_yr", "taxes_calc", 
    matches("area")) %>% 
  # formatting for map
  mutate(across(c("assessed_adj",
                "assessed_adj_orig",
                "taxes_yr",
                "taxes_calc",
                "bldgarea"), ~if_else(is.na(.) | is.infinite(.), "", prettyNum(., big.mark = ",")))) %>% 
  mutate(across(c("address_form",
                "ownername"), ~if_else(is.na(.) | is.infinite(.), "", .))) %>% 
  mutate(across(c("assessed_adj",
                "assessed_adj_orig",
                "taxes_yr",
                "taxes_calc"), ~if_else(. != "", str_c("$",.), "")))
nyu_assess_geo
```
## Attempting to fix missing geometries
* Attempt at geocoding
```{r}
geoms2geocode <- nyu_assess_geo %>% 
  filter(st_is_empty(geometry)) %>% 
  as_tibble() %>% 
  select(-"geometry")

geoms_geocoded <- geoms2geocode %>% 
  geo_search_data(address_form) %>% 
  rename(bbl_geo = bbl) %>%
  select(c("bbl_geo", "no_results")) %>% 
  cbind(geoms2geocode) %>% 
  filter(no_results == F) %>% 
  bind_rows(
    geoms2geocode %>% 
  filter(is.na(address_form))
  )  %>% 
  # need to manually code this bbl
  mutate(bbl = case_when(bbl == "2057530140" ~ "2057520121",
                         bbl != bbl_geo ~ bbl_geo,
                         TRUE ~ bbl))
geoms_geocoded

copy_to(con, geoms_geocoded,
  name = "geoms_geocoded", overwrite = T,
  temporary = T
)
fix_geom <- st_read(con, query = "SELECT g.*, m21.geom_poly as geom1,
                                          m2.geom_poly as geom2 FROM geoms_geocoded as g
                        LEFT JOIN mappluto_21v3 as m21 on (g.bbl = m21.bbl)
                        LEFT JOIN mappluto_02b as m2 on (g.bbl = m2.bbl)")
fix_geom %>% 
  as_tibble() %>% 
  st_set_geometry("geom2") %>% 
  filter(st_is_empty(geom2)) %>% 
  nrow() %>% 
  equals(0) %>% 
  stopifnot()
fix_geom


fix_geom_filt <- fix_geom %>% 
  as_tibble() %>% 
  rename(geometry = geom2) %>% 
  st_set_geometry("geometry") %>%
  select(-c("geom1")) %>% 
  group_by(bbl, year) %>% 
  filter(row_number() == 1) %>% 
  st_transform(4326)
fix_geom_filt
```

```{r}
nyu_assess_fix <- nyu_assess_geo %>% 
  filter(!st_is_empty(geometry)) %>% 
  bind_rows(fix_geom_filt) %>% 
  unique() 

nyu_assess_fix %>% 
  filter(st_is_empty(geometry)) %>% 
  nrow() %>% 
  equals(0) %>% 
  stopifnot()

nyu_assess_fix %>% 
  tablist_qc(year)
  
nyu_assess_fix %>% 
  # make sure not to include earlier years
  # new decision: only show after 2002
  # there are some issues with condos showing up as the whole building
  filter(year >= 2002) %>% 
  st_write(
    dsn = path(
      "04_WebMap",
      "data",
      "nyu_clean.geojson"
    ), driver = "GeoJSON",
    append = F,
    delete_dsn = T
  )
```

# Analysis of taxes
```{r}
nyu_assess_value %>% 
  as_tibble() %>% 
  filter(year >= 2002) %>% 
  select(c("taxes_yr", "year")) %>% 
  unique() %>% 
  verify_isid(c("year")) %>% 
  mutate(taxes_yr = taxes_yr / 1000) %>% 
  ggplot(aes(year, taxes_yr)) +
  geom_line(color = colors[1]) +
  theme_classic() +
  scale_y_continuous(limits = c(0, 200000)) +
  labs(x = "Year",
      y = "Estimated taxes in thousands") +
  scale_x_continuous(limits = c(2001, 2021),
                     breaks = seq(2001, 2021, 2))

ggsave(path(fig_path, "estimated_taxes.pdf"))

# cumulative taxes (over time)
nyu_assess_value %>% 
  as_tibble() %>% 
  filter(year >= 2002) %>% 
  select(c("taxes_yr", "year")) %>% 
  unique() %>% 
  verify_isid(c("year")) %>% 
  arrange(year) %>% 
  mutate(taxes_cum = cumsum(taxes_yr) / 1000) %>% 
  ggplot(aes(year, taxes_cum)) +
  geom_line(color = colors[1]) +
  theme_classic() +
  scale_y_continuous(limits = c(0, 3000000),
                     breaks = c(0, 1000000, 2000000, 3000000),
                     labels = c("$0", "$1,000,000", "$2,000,000", "$3,000,000")) +
  labs(x = "Year",
      y = "Cumulative estimated taxes in thousands") +
  scale_x_continuous(limits = c(2001, 2021),
                     breaks = seq(2001, 2021, 2))

ggsave(path(fig_path, "estimated_taxes_cum.pdf"))
```


# Analysis of number of buildings
```{r}
nyu_poly  %>% 
  unique() %>% 
  mutate(borough = str_sub(cd, 1, 1),
         borough = recode(borough, "3" = "Brooklyn", "1" = "Manhattan",
                          "2" = "Bronx", "5" = "Staten Island",
                          "4" = "Queens")) %>% 
  filter(borough == "Bronx") %>% 
  group_by(year) %>% 
  summarize(n = n())

nyu_poly %>% 
  mutate(borough = recode(borough, "3" = "Brooklyn", "1" = "Manhattan",
                          "2" = "Bronx", "5" = "Staten Island",
                          "4" = "Queens"))  %>% 
  #filter(borough != "Bronx") %>% 
  mutate(borough = factor(borough, levels = c("Manhattan", "Brooklyn",
                                              "Bronx"),
                          ordered = T)) %>% 
  verify(!is.na(borough)) %>% 
  group_by(year, borough) %>% 
  summarize(count = n_distinct(bbl)) %>% 
  ggplot() +
  geom_area(aes(year, count, fill = borough)) +
  scale_y_continuous(limits = c(0, 170)) +
  scale_x_continuous(limits = c(2001, 2021),
                     breaks = seq(2001, 2021, 2)) +
  theme_classic() +
  labs(x = "Year",
       y = "Number of buildings") +
  scale_fill_manual(values = colors)

ggsave(path(fig_path, "num_bld_time.pdf"))
```

# Analysis of square footage (building or floor square footage)
```{r}
nyu_area <- nyu_poly %>% 
  mutate(bldgarea = if_else(is.na(bldgarea), floorarea, as.double(bldgarea))) %>% 
  as_tibble() %>% 
  verify(!is.na(bldgarea)) %>% 
  group_by(borough, year) %>%
  summarize(bldgarea_total_boro = sum(bldgarea, na.rm = T))  %>% 
  group_by(year) %>%
  mutate(bldgarea_total = sum(bldgarea_total_boro, na.rm = T)) %>% 
  rowwise() %>%
  mutate(bldgarea_share = round(bldgarea_total_boro * 100 / bldgarea_total)) %>% 
  ungroup() %>% 
  mutate(across(c("bldgarea_total", "bldgarea_total_boro"), ~./1000)) %>% 
  mutate(borough = recode(borough, "3" = "Brooklyn", "1" = "Manhattan",
                          "2" = "Bronx", "5" = "Staten Island",
                          "4" = "Queens"))

# stacked area graph
nyu_area %>% 
  mutate(borough = factor(borough, levels = c("Manhattan", 
                                              "Brooklyn",
                                              "Bronx"),
                          ordered = T)) %>% 
  select(c("borough", "year", "bldgarea_total_boro")) %>% 
  unique() %>% 
  rename(bldgarea_total = bldgarea_total_boro) %>% 
  # bind_rows(
  #   nyu_area %>% 
  #   select(c( "year", "bldgarea_total")) %>% 
  #   unique() %>% 
  #   mutate(borough = "Total")) %>% 
  ggplot() +
  geom_area(aes(year, bldgarea_total, fill = borough)) +
  theme_classic() +
  labs(x = "Year",
       y = "Square feet in thousands") +
  scale_fill_manual(values = colors) +
  scale_x_continuous(limits = c(1997, 2021),
                     breaks = seq(1997, 2021, 2))

ggsave(path(fig_path, "area_1.pdf"))
```

# Repeat the above analyses with 
## Analysis of number of buildings
```{r}
rpad_allyrs  %>% 
  verify_isid(c("bbl", "year")) %>% 
  mutate(borough = str_sub(cd, 1, 1),
         borough = recode(borough, "3" = "Brooklyn", "1" = "Manhattan",
                          "2" = "Bronx", "5" = "Staten Island",
                          "4" = "Queens")) %>% 
  filter(borough == "Bronx") %>% 
  nrow()

rpad_allyrs %>% 
  mutate(borough = str_sub(cd, 1, 1),
         borough = recode(borough, "3" = "Brooklyn", "1" = "Manhattan",
                          "2" = "Bronx", "5" = "Staten Island",
                          "4" = "Queens"))  %>% 
  #filter(borough != "Bronx") %>% 
  mutate(borough = factor(borough, levels = c("Manhattan", "Brooklyn",
                                              "Bronx"),
                          ordered = T)) %>% 
  verify(!is.na(borough)) %>% 
  group_by(year, borough) %>% 
  summarize(count = n_distinct(bbl)) %>% 
  ggplot() +
  geom_area(aes(year, count, fill = borough)) +
  scale_y_continuous(limits = c(0, 300)) +
  scale_x_continuous(limits = c(1997, 2021),
                     breaks = seq(1997, 2021, 2)) +
  theme_classic() +
  labs(x = "Year",
       y = "Number of buildings") +
  scale_fill_manual(values = colors)

ggsave(path(fig_path, "num_bld_time_rpad.pdf"))
```

## Analysis of square footage (building or floor square footage)
```{r}
nyu_area_rpad <- rpad_allyrs %>% 
  verify_isid(c("bbl", "year")) %>%
  verify(!is.na(bldgarea))  %>% 
  mutate(borough = str_sub(cd, 1, 1),
         borough = recode(borough, "3" = "Brooklyn", "1" = "Manhattan",
                          "2" = "Bronx", "5" = "Staten Island",
                          "4" = "Queens")) %>% 
  group_by(borough, year) %>%
  summarize(bldgarea_total_boro = sum(bldgarea, na.rm = T))  %>% 
  group_by(year) %>%
  mutate(bldgarea_total = sum(bldgarea_total_boro, na.rm = T)) %>% 
  rowwise() %>%
  mutate(bldgarea_share = round(bldgarea_total_boro * 100 / bldgarea_total)) %>% 
  ungroup() %>% 
  mutate(across(c("bldgarea_total", "bldgarea_total_boro"), ~./1000))

# stacked area graph
nyu_area_rpad %>% 
  #filter(borough == "Manhattan" | borough == "Brooklyn") %>% 
  mutate(borough = factor(borough, levels = c("Manhattan", "Brooklyn",
                                              "Bronx"),
                          ordered = T)) %>% 
  select(c("borough", "year", "bldgarea_total_boro")) %>% 
  unique() %>% 
  rename(bldgarea_total = bldgarea_total_boro) %>% 
  # bind_rows(
  #   nyu_area %>% 
  #   select(c( "year", "bldgarea_total")) %>% 
  #   unique() %>% 
  #   mutate(borough = "Total")) %>% 
  ggplot() +
  geom_area(aes(year, bldgarea_total, fill = borough)) +
  theme_classic() +
  labs(x = "Year",
       y = "Square feet in thousands") +
  scale_fill_manual(values = colors) +
  scale_x_continuous(limits = c(1997, 2021),
                     breaks = seq(1997, 2021, 2))

ggsave(path(fig_path, "area_rpad.pdf"))
```

## Comparison of 2002 vs. 2021 via maps
```{r}
nyu_pts  %>% 
  filter(year == 2002 | year == 2021)  %>% 
  as_tibble() %>% 
  verify(!is.na(cd)) %>% 
  mutate(boro_cd = as.character(cd)) %>% 
  group_by(boro_cd, year) %>% 
  summarize(n_bld = n_distinct(bbl)) %>% 
  inner_join(cds, ., by = "boro_cd")  %>% 
  ggplot() +
  geom_sf(data = boro, fill = NA, color = "white", size = 0.01) +
  geom_sf(aes(fill = n_bld), color = NA) +
  geom_sf_text(aes(label = n_bld), size = 2, color = "white") +
  theme_void()  +
  theme(
    panel.border = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    plot.background = element_rect(fill = '#A5A5A5', color = NA),
    panel.background = element_rect(fill = '#A5A5A5', color = NA)
  ) +
  scale_fill_gradient(
    low = "#CCB7E5",
    high = alt_purple) +
  facet_grid(cols = vars(year))

ggsave(path(fig_path, "nyu_pts_year_comp.pdf"), width = 9, height = 9)

rpad_allyrs %>% 
  verify_isid(c("bbl", "year")) %>% 
  filter(year == 2002 | year == 2021)  %>% 
  as_tibble() %>% 
  verify(!is.na(cd)) %>% 
  mutate(boro_cd = as.character(cd)) %>% 
  group_by(boro_cd, year) %>% 
  summarize(n_bld = n_distinct(bbl)) %>%
  inner_join(cds, ., by = "boro_cd") %>% 
  ggplot() +
  geom_sf(data = boro, fill = NA, color = "white", size = 0.01) +
  geom_sf(aes(fill = n_bld), color = NA) +
  geom_sf_text(aes(label = n_bld), size = 2, color = "white") +
  theme_void()  +
  theme(
    panel.border = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    plot.background = element_rect(fill = '#A5A5A5', color = NA),
    panel.background = element_rect(fill = '#A5A5A5', color = NA)
  ) +
  scale_fill_gradient(
    low = "#CCB7E5",
    high = alt_purple) +
  facet_grid(cols = vars(year))

ggsave(path(fig_path, "nyu_rpad_year_comp.pdf"), width = 9, height = 9)

rpad_allyrs %>% 
  filter(str_sub(cd, 1, 1) == 3) %>% 
  filter(year == 2021) %>% 
  select("cd", everything()) %>% 
  arrange(cd)

rpad_allyrs %>% 
  filter(cd == 106)

rpad_allyrs %>% 
  filter(str_detect(str_name, "MAIN")) %>% 
  select(c("year"), everything())

rpad_allyrs %>% 
  tablist_qc(exmtcl)

rpad_allyrs %>% 
  #filter(bbl != bbl_condo) %>%
  arrange(-year) %>% 
  select(c("year", "ownername"), everything())
``` 



